---
title: "Shiny"
author: "Tash Ramsden"
date: "12/08/2021"
output: html_document
---

## Shiny Uses

* Interactive data visualisation for presentations and websites
* Sharing results with collaborators
* Communicating science in an accessible way
* Bridging the gap between R users and non-R users
* See some examples [here](https://shiny.rstudio.com/gallery/)


## Set-up

Run the following to set up Shiny in RStudio:

* `install.packages("shiny")`
* `install.packages("rsconnect")  # For publishing apps online`
* `install.packages("agridat")  # For the dataset in today's tutorial`


## Shiny app file structure

Create the Shiny app file:

* *File/New File/Shiny Web App...*
* Application type: *Single File (app.R)*
* Name of app assigned to dir (not the app script file - called app.R)
* (Shiny apps can be made with 2 files: ui.R and server.R as well)

Inside the app directory:

* Create a folder called **Data** - to hold data!
* Create a folder called **www** - to hold any images/other web elements


## app.R Layout

**A basic app.R contains these 5 parts:**

1. A section at the top loading packages needed for the app 
    + **`library(shiny)`**
    + `shiny` is required
2. A section to load any data
    + **`Barley <- as.data.frame(beaven.barley)`**
3. An object called `ui` containing info about app layout in the browser
    + **`ui <- fluidPage()`**
    + `fluidpage()` defines a layout that resizes according to window size (is responsive)
    + All the app code will be placed inside the brackets
4. A `server` object containing info about app computation, plotting, tables, maps etc
    + **`server <- function(input, output) {}`**
    + All the app code will be placed inside the curly brackets
5. A command to run the app
    + **`shinyApp(ui = ui, server = server)`**
    + Should be at the top of `app.R`
    + Tells shiny that the user interface comes from `ui` and server info (data, plots, tables, etc) comes from `server`


## Shiny App Layout

Shiny apps are structured using panels which can contain text, widgets, plots, tables, maps, images, etc.

* Some [layout examples](https://shiny.rstudio.com/articles/layout-guide.html)
* Most basic: fluidRow() and column() to manually create grids
* sidebarLayout() creates large panel and smaller inset side panel


## Basic Syntax for Creating a Shiny App


### Define a layout within `ui`, e.g:

```{r, eval=FALSE}
ui <- 
    fluidPage(
        titlePanel(""),
        sidebarLayout(
            sidebarPanel(),
            mainPanel()
        )
    )
```

* `titlePanel()` indicates that we would like a separate panel at the top of the page in which we can put the title.
* `sidebarLayout()` indicates that we want our Shiny app to have the sidebar layout, one of many layouts we saw above. Within sidebarLayout we have:
* `sidebarPanel()` indicates that we want a sidebar panel included in our app. Sidebar panels often contain input widgets like sliders, text input boxes, radio buttons etc.
* `mainPanel()` indicates that we want a larger main panel. Main panels often contain the output of the app, whether it is a table, map, plot or something else.


### Input Widgets

Once the basic structure is defined, add inputs and outputs to `ui`, e.g:

```{r, eval=FALSE}
ui <- fluidPage(
  titlePanel("Barley Yield"),
  sidebarLayout(
    sidebarPanel(
      selectInput(inputId = "gen",  # Give the input a name "genotype"
      label = "1. Select genotype",  # Give the input a label to be displayed in the app
      choices = c("A" = "a","B" = "b","C" = "c","D" = "d","E" = "e","F" = "f","G" = "g","H" = "h"), selected = "a"),  # Create the choices that can be selected. e.g. Display "A" and link to value "a"
      # Or: choices = unique(Barley$gen)
    selectInput(inputId = "colour", 
      label = "2. Select histogram colour", 
      choices = c("blue","green","red","purple","grey"), selected = "grey"),
    sliderInput(inputId = "bin", 
      label = "3. Select number of histogram bins", 
      min=1, max=25, value= c(10)),
    textInput(inputId = "text", 
      label = "4. Enter some text to be displayed", "")
    ),
  mainPanel()
  )
)
```

* For help with understanding the arguments for input widgets use the help function `?`, e.g. `?textInput`
* Other widgets, e.g. radio buttons, action buttons, etc


## Running a Shiny App

**Click *Run App* ** or **`Ctrl + Alt + R`**

Stop the app by pressing *Stop* button or `Esc`


## Output

Without outputs the app doesn't tell us anything - outputs can be plots, tables, maps, etc

Create outputs by placing code inside the curly brackets `{}` in `server`, e.g:

```{r, eval=FALSE}
server <- function(input, output) {
  output$plot <- renderPlot(ggplot(Barley, aes(x = yield)) +  # Create object called `output$plot` with a ggplot inside it
  geom_histogram(bins = 7,  # Add a histogram to the plot
    fill = "grey",  # Make the fill colour grey
    data = Barley,  # Use data from `Barley`
    colour = "black")  # Outline the bins in black
  )          
}
```

* Created an object called output$plot and used `renderPlot()` to wrap a `ggplot()` command.


## Reactive output

Need to link the input widgets to the output object so that it is interactive.

1. Update `server` with new reactive output arguments:

```{r, eval=FALSE}

server <- function(input, output) {
  output$plot <- renderPlot(ggplot(Barley, aes(x = yield)) +
  geom_histogram(bins = 7,
      fill = "grey",
      data = Barley[Barley$gen == input$gen,],
      colour = "black")
  )                                                       
}

```

* `data = Barley[Barley$gen == input$gen,]` means only use data where value in column `gen` is equal to value given by `input$gen`, all rows selected (just filter columns)

2. Update histogram colour based on value of colour input:

```{r, eval=FALSE}

      fill = input$colour

```

3. Select number of bins in histogram:

```{r, eval=FALSE}

      bins = input$bin

```

4. Finally, create a table output showing some summary stats for the selected parameters.

* Create new output object `output$table` in `server` and use `renderTable()` to create the table made using dplyr `summarise()`:


```{r, eval=FALSE}

server <- function(input, output) {
  output$myhist <- renderPlot(ggplot(Barley, aes(x = yield))   + 
  	geom_histogram(bins = input$bin, fill = input$col,   group=input$gen, 
  		data=Barley[Barley$gen == input$gen,],
    		colour = "black"))
  
  output$mytext <- renderText(input$text)
  
  output$mytable <- renderTable(Barley %>%
  filter(gen == input$gen) %>%
  summarise("Mean" = mean(yield), 
    	"Median" = median(yield),
    	"STDEV" = sd(yield), 
    	"Min" = min(yield),
    	"Max" = max(yield)))
}

```


## Displaying Output

Need to make outputs appear on app in `mainPanel` - add to `ui` object inside `mainPanel`:

```{r, eval=FALSE}

    mainPanel(
      plotOutput("plot"),
      tableOutput("table"),
      textOutput("text")
    )

```

*At this point, preview the app - it's all working!*

## Additional elements

### HTML

Like a normal HTML webpage, HTML tags can be added to shiny apps, e.g:

| HTML  | Shiny                                 | Function                                |
|-------|---------------------------------------|-----------------------------------------|
|`<div>`|`tags$div()`                           |Defines block with consistent formatting |
|`<br>` |`tags$br()`                            |Inserts break                            |
|`<hr>` |`tags$hr()`                            |Inserts horizontal line                  |
|`<p>`  |`tags$p()`                             |Creates paragraph of text                |
|`<a>`  |`tags$a(href="LINK", "displayed text")`|Creates clickable link                   |


* Find list of all HTML tags using `shiny::tags`

* Some tags can conflict with other functions so should always state the source by using `tags$`

* Tags can be stacked as in HTML:

```{r, eval=FALSE}

tags$div(style="color:red",
  tags$p("Visit us at:"),
  tags$a(href = "https://ourcodingclub.github.io", "Coding Club")
  )

```

* Above creates a block of text that is red, the block contains a paragraph saying "Visit us at:", and a link to the Coding Club website (add above code into `mainPanel()` to view!)

For more Shiny arguments see [RStudio wiki](https://shiny.rstudio.com/articles/tag-glossary.html)


## Exporting a finished app

### As a GitHub repository

It's easy to share Shiny apps with people who have RStudio:

* Easiest way is to send app.R (w any data, resources) in a zip file to be unzipped by recipient and run through R:

* Can quickly share over the internet using Github:

  + Create repo and upload everything from app folder (including any `Data` and `www` folders)
  
  + Add README.md 
  
  + Give recipient Github username and name of app repo
  
  + Get them to run `runGithub()` in R: `runGithub(repo = "repo_name", username = "user_name")`

* Alternatively (if recipient doesn't know how Github works) upload folder as `.zip` file to Github (or any other file-hosting service) and they can use `runUrl()` if given the url of the zipfile, eg: `runUrl("https://github.com/rstudio/shiny_example/archive/master.zip")`


### As a shinyapps.io app

Shiny apps can be hosted on [www.shinyapps.io](www.shinyapps.io), a web hosting
platform by RStudio for SHiny apps.

* Sign up on their website
* Then go to [www.shinyapps.io/admin/#/tokens](www.shinyapps.io/admin/#/tokens), 
click Show secret and copy the rsconnect account info
* Run this code in R to link shinyapps.io with R Studio
* To upload your app, open your app.R and click the publish button. Select a 
name for your app (no spaces) and click Publish.
* The app can then be launched by anyone with the URL (can be found by going ot shinyapps.io and opening app info from dashboard)

* To embed an app that is hosted by `shinyapps.io` in your own website you can put it in an `iframe`, replacing the URL with your own app URL and altering the style arguments to your own desire:

```{r eval=FALSE}

<iframe src="https://johngodlee.shinyapps.io/barley_yield_exploration/" style="border:none;width:1000px;height:500px;"></iframe>

```

